---
title: "Comparison of Harbour Porpoise clicks"
author: "Lina Fliess"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
install.packages("psych")
```

```{r librarys, echo=FALSE}
library(readxl)
library(ggplot2)
library(car)
library(psych)
```

```{r data DE, echo=FALSE}
Ref_DE_Aschau <- read_excel("AS_78 2023 02 28 FPOD_6889_train_details_filtered.xlsx")
Ref_DE_BoEck <- read_excel("BoEck 2023 02 07 FPOD_6877_train_details.xlsx")
Ref_DE_Langholz <- read_excel("LA_77 2023 02 28 FPOD_6888_train_details_filtered.xlsx")
#bind all dataframes from german ref stations
Ref_DE <- rbind(Ref_DE_Aschau, Ref_DE_BoEck, Ref_DE_Langholz)
```

```{r data DK, echo=FALSE}     
Ref_DK_6621 <- read_excel("2023 02 21 FPOD_6621 file0_train_details.xlsx")
Ref_DK_6623 <- read_excel("2023 02 21  2023 02 21 FPOD_6623 file0 train details.xlsx")
Ref_DK_6625 <- read_excel("20230221 2023 02 21 FPOD_6625 file0 train details.xlsx")
#bind all dataframes from danish ref stations
Ref_DK <- rbind(Ref_DK_6621, Ref_DK_6623, Ref_DK_6625)
```

```{r groups, echo=FALSE}   
Ref_DE$group <- 'Dataset DE'
Ref_DK$group <- 'Dataset DK'
Ref_both <- rbind(Ref_DE, Ref_DK)
```

## <span style="color:blue">Comparison of median frequency of a whole train</span>###

## Hypothesis

H_0: There is no difference in the distribution of the median frequency (whole train)
between the german and danish harbour porpoise clicks

## Exploratory Analysis

```{r summary}
summary(Ref_DE$medianKHz)
summary(Ref_DK$medianKHz)
```

Visualization/Plots:

```{r histogram, echo=FALSE}
ggplot() +
  geom_histogram(data = Ref_DE, aes(x = medianKHz), bins = 50, fill = 'blue', alpha = 0.5) +
  geom_histogram(data = Ref_DK, aes(x = medianKHz), bins = 50, fill = 'orange', alpha = 0.5) +
  labs(title = "Mean Frequency Distribution", x = "Mean Frequency", y = "Count") +
  theme_minimal() +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +
  theme(legend.position = "top") 
```

```{r density plot, echo=FALSE}
ggplot() +
  geom_density(data = Ref_DE, aes(x = medianKHz), fill = 'blue', alpha = 0.5) +
  geom_density(data = Ref_DK, aes(x = medianKHz), fill = 'orange', alpha = 0.5) +
  labs(title = "Density Plot of Mean Frequency", x = "Mean Frequency", y = "Density") +
  theme_minimal() +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +
  theme(legend.position = "top") 
```

```{r boxplot, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = medianKHz, fill = group)) +
  geom_boxplot() +
  labs(title = "Boxplot of Mean Frequency", x = "Group", y = "Mean Frequency") +
  theme_minimal()
```

```{r scatter plot, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = medianKHz, color = group)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(title = "Scatter Plot of Mean Frequency", x = "Group", y = "Mean Frequency") +
  theme_minimal()
```

## Check Requirements for tests

### Normality of Distribution

- size of Dataframe too large for Shapiro-Wilk Test

- using Q-Q-Plots to get a visual idea and Kolmogorov-Smirnov Test as an alternative

Median Frequency DE:

```{r normalityDE, echo=FALSE}
#checking if data follows normal distribution (Shapiro-Wilk,  Kolmogorov-Smirnov or Q-Q-Plots)
qqnorm(Ref_DE$medianKHz)
qqline(Ref_DE$medianKHz)
ks.test(Ref_DE$medianKHz, "pnorm", mean = mean(Ref_DE$medianKHz), sd = sd(Ref_DE$medianKHz))
# no normal distribution, significant difference between sample and normal distribution (ks test)
```

Median Frequency DK:

```{r normalityDK, echo=FALSE}
qqnorm(Ref_DK$medianKHz)
qqline(Ref_DK$medianKHz)
ks.test(Ref_DK$medianKHz, "pnorm", mean = mean(Ref_DK$medianKHz), sd = sd(Ref_DK$medianKHz))
#very small p-value (< 2.2e-16) suggests that the sample distribution significantly differs from the normal distribution
```

No noticeable normal distribution in Q-Q-Plot which is underlined by the result of the Kolmogorov-Smirnov-Test. <br>
The small p-value (<2.2e-16) suggest that the distribution significantly differs from normal distribution. <br>

Also checking for homogeneity of variances (Levenes Test):

```{r leveneTest, echo=FALSE}
leveneTest(medianKHz ~ as.factor(c(rep(1, nrow(Ref_DE)), rep(2, nrow(Ref_DK)))), data = data.frame(medianKHz = c(Ref_DE$medianKHz, Ref_DK$medianKHz)))
#p-value (0.004727) < 0.05 --> statistically significant difference in variances between the groups
#null hypothesis of homogeneity of variances rejected
```

p-value (0.004727) < 0.05 <br>
--> statistically significant difference in variances between the groups <br>
Therefore the null hypothesis of homogeneity of variances is rejected. <br>

## Statistical Test (Wilcoxon)

```{r test, echo=FALSE}
#t.test(Ref_DE$medianKHz, Ref_DK$medianKHz, var.equal = FALSE)
wilcox.test(Ref_DE$medianKHz, Ref_DK$medianKHz)
```

p (<2.2e-16) < 0.05 <br>
--> there is a significant difference between the datasets for the criterion median frequency

## <span style="color:blue">Comparison of average frequency of last Cycle</span>##

## Hypothesis

H_0: There is no difference in the distribution of the average frequency of the last cycle between the german and danish harbour porpoise clicks

## Exploratory Analysis

```{r summary 2}
summary(Ref_DE$avEndF)
summary(Ref_DK$avEndF)
```

Visualization/Plots:

```{r histogram 2, echo=FALSE}
ggplot() +
  geom_histogram(data = Ref_DE, aes(x = avEndF), bins = 50, fill = 'blue', alpha = 0.5) +
  geom_histogram(data = Ref_DK, aes(x = avEndF), bins = 50, fill = 'orange', alpha = 0.5) +
  labs(title = "Average Frequency of last Cycle", x = "Average Frequency", y = "Count") +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +   
  theme_minimal() +
  theme(legend.position = "top") 
```

```{r density plot 2, echo=FALSE}
ggplot() +
  geom_density(data = Ref_DE, aes(x = avEndF), fill = 'blue', alpha = 0.5) +
  geom_density(data = Ref_DK, aes(x = avEndF), fill = 'orange', alpha = 0.5) +
  labs(title = "Density Plot of Average Frequency of last Cycle", x = "Average Frequency", y = "Density") +
  theme_minimal() +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +
  theme(legend.position = "top") 
```

```{r boxplot 2, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = avEndF, fill = group)) +
  geom_boxplot() +
  labs(title = "Boxplot of Average Frequency of last Cycle", x = "Group", y = "Average Frequency") +
  theme_minimal()
```

```{r scatter plot 2, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = avEndF, color = group)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(title = "Scatter Plot of Average Frequency of last Cycle", x = "Group", y = "Average Frequency") +
  theme_minimal()
```

## Check Requirements for tests

### Normality of Distribution

Average end frequency DE:

```{r normalityDE 2, echo=FALSE}
#checking if data follows normal distribution (Shapiro-Wilk (max. 5000 samples),  Kolmogorov-Smirnov or Q-Q-Plots)
qqnorm(Ref_DE$avEndF)
qqline(Ref_DE$avEndF)
ks.test(Ref_DE$avEndF, "pnorm", mean = mean(Ref_DE$avEndF), sd = sd(Ref_DE$avEndF))
# no normal distribution, significant difference between sample and normal distribution (ks test)
```

Average end frequency DK:

```{r normalityDK 2, echo=FALSE}
qqnorm(Ref_DK$avEndF)
qqline(Ref_DK$avEndF)
ks.test(Ref_DK$avEndF, "pnorm", mean = mean(Ref_DK$avEndF), sd = sd(Ref_DK$avEndF))
#very small p-value (< 2.2e-16) suggests that the sample distribution significantly differs from the normal distribution
```

No noticeable normal distribution in Q-Q-Plot which is underlined by the result of the Kolmogorov-Smirnov-Test.<br>
The small p-value (<2.2e-16) suggest that the distribution significantly differs from normal distribution.
<br>
Also checking for homogeneity of variances (Levenes Test):

```{r leveneTest 2, echo=FALSE}
leveneTest(avEndF ~ as.factor(c(rep(1, nrow(Ref_DE)), rep(2, nrow(Ref_DK)))), data = data.frame(avEndF = c(Ref_DE$avEndF, Ref_DK$avEndF)))
#p-value (<2.2e-16) < 0.05 --> statistically significant difference in variances between the groups
#null hypothesis of homogeneity of variances rejected
```

p-value (<2.2e-16) < 0.05 <br>
--> statistically significant difference in variances between the groups <br>
Meaning the null hypothesis of homogeneity of variances is rejected.

## Statistical Test

```{r test 2, echo=FALSE}
#t.test(Ref_DE$avEndF, Ref_DK$avEndF, var.equal = FALSE)
wilcox.test(Ref_DE$avEndF, Ref_DK$avEndF)
#p (<2.2e-16) < 0.05 --> there is a significant difference between the datasets for the median frequency of the last cycle of a train
```

p (<2.2e-16) < 0.05 <br>
--> there is a significant difference between the datasets for the average frequency of the last cycle

## <span style="color:blue">Comparison of average sound Pressure Level</span>###

## Hypothesis

H_0: There is no difference in the distribution of the average sound pressure level between the german and danish harbour porpoise clicks

## Exploratory Analysis

```{r summary 3}
summary(Ref_DE$avSPL)
summary(Ref_DK$avSPL)
```

Visualization/Plots:

```{r histogram 3, echo=FALSE}
ggplot() +
  geom_histogram(data = Ref_DE, aes(x = avSPL), bins = 50, fill = 'blue', alpha = 0.5) +
  geom_histogram(data = Ref_DK, aes(x = avSPL), bins = 50, fill = 'orange', alpha = 0.5) +
  labs(title = "Average Sound Pressure Level", x = "Average SPL", y = "Count") +
  theme_minimal() +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +
  theme(legend.position = "top")
```

```{r density plot 3, echo=FALSE}
ggplot() +
  geom_density(data = Ref_DE, aes(x = avSPL), fill = 'blue', alpha = 0.5) +
  geom_density(data = Ref_DK, aes(x = avSPL), fill = 'orange', alpha = 0.5) +
  labs(title = "Density Plot of Average Sound Pressure Level", x = "Average SPL", y = "Density") +
  theme_minimal() +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +
  theme(legend.position = "top")
```

```{r boxplot 3, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = avSPL, fill = group)) +
  geom_boxplot() +
  labs(title = "Boxplot of Average Sound Pressure Level", x = "Group", y = "Average SPL") +
  theme_minimal()
```

```{r scatter plot 3, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = avSPL, color = group)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(title = "Scatter Plot of Average Sound Pressure Level", x = "Group", y = "Average SPL") +
  theme_minimal()
```

## Check Requirements

### Normality of Distribution

Average Sound Pressure Level DE:
```{r normalityDE 3, echo=FALSE}
qqnorm(Ref_DE$avSPL)
qqline(Ref_DE$avSPL)
ks.test(Ref_DE$avSPL, "pnorm", mean = mean(Ref_DE$avSPL), sd = sd(Ref_DE$avSPL))
# no normal distribution, significant difference between sample and normal distribution (ks test)
```
Average Sound Pressure Level DK:
```{r normalityDK 3, echo=FALSE}
qqnorm(Ref_DK$avSPL)
qqline(Ref_DK$avSPL)
ks.test(Ref_DK$avSPL, "pnorm", mean = mean(Ref_DK$avSPL), sd = sd(Ref_DK$avSPL))
#very small p-value (< 2.2e-16) suggests that the sample distribution significantly differs from the normal distribution
```

No noticeable normal distribution in Q-Q-Plot which is underlined by the result of the Kolmogorov-Smirnov-Test. <br>
The small p-value (<2.2e-16) suggest that the distribution significantly differs from normal distribution. <br>

Also checking for homogeneity of variances (Levenes Test):

```{r leveneTest 3, echo=FALSE}
leveneTest(avSPL ~ as.factor(c(rep(1, nrow(Ref_DE)), rep(2, nrow(Ref_DK)))), data = data.frame(avSPL = c(Ref_DE$avSPL, Ref_DK$avSPL)))
#p-value (<2.2e-16) < 0.05 --> statistically significant difference in variances between the groups
#null hypothesis of homogeneity of variances rejected
```

p-value (<2.2e-16) < 0.05 <br>
--> statistically significant difference in variances between the groups <br>
Meaning the null hypothesis of homogeneity of variances is rejected. <br>

## Statistical Test

```{r test 3, echo=FALSE}
#t.test(Ref_DE$avSPL, Ref_DK$avSPL, var.equal = FALSE)
wilcox.test(Ref_DE$avSPL, Ref_DK$avSPL)
#p (<2.2e-16) < 0.05 
#--> there is a significant difference between the datasets for the average sound pressure level
```

p (<2.2e-16) < 0.05 <br>
--> there is a significant difference between the datasets for the average sound pressure level <br>
Null Hypothesis rejected.

## <span style="color:blue">Comparison of average Train Duration (micro seconds) </span>###

## Hypothesis

H_0: There is no difference in the distribution of the average Train Duration between the german and danish harbour porpoise clicks

## Exploratory Analysis

```{r summary 4}
summary(Ref_DE$TrDur_us)
summary(Ref_DK$TrDur_us)
```

Visualization/Plots:

```{r histogram 4, echo=FALSE}
ggplot() +
  geom_histogram(data = Ref_DE, aes(x = TrDur_us), bins = 50, fill = 'blue', alpha = 0.5) +
  geom_histogram(data = Ref_DK, aes(x = TrDur_us), bins = 50, fill = 'orange', alpha = 0.5) +
  labs(title = "Train Duration Distribution", x = "Average Train Duration", y = "Count") +
  theme_minimal() +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +
  theme(legend.position = "top")
```

```{r density plot 4, echo=FALSE}
ggplot() +
  geom_density(data = Ref_DE, aes(x = TrDur_us), fill = 'blue', alpha = 0.5) +
  geom_density(data = Ref_DK, aes(x = TrDur_us), fill = 'orange', alpha = 0.5) +
  labs(title = "Density Plot of Average Train Duration", x = "Average Train Duration", y = "Density") +
  theme_minimal() +
  scale_fill_manual(name = "Dataset", values = c("DE" = "blue", "DK" = "orange")) +
  theme(legend.position = "top")
```

```{r boxplot 4, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = TrDur_us, fill = group)) +
  geom_boxplot() +
  labs(title = "Boxplot of Average Train Duration", x = "Group", y = "Average Train Duration") +
  theme_minimal()
```

```{r scatter plot 4, echo=FALSE}
ggplot(Ref_both, aes(x = group, y = TrDur_us, color = group)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(title = "Scatter Plot of Average Train Duration", x = "Group", y = "Average Train Duration") +
  theme_minimal()
```

## Check Requirements

### Normality of Distribution 

Train Duration DE:

```{r normalityDE 4, echo=FALSE}
qqnorm(Ref_DE$TrDur_us)
qqline(Ref_DE$TrDur_us)
ks.test(Ref_DE$TrDur_us, "pnorm", mean = mean(Ref_DE$TrDur_us), sd = sd(Ref_DE$TrDur_us))
# no normal distribution, significant difference between sample and normal distribution (ks test)
```

Train Duration DK:

```{r normalityDK 4, echo=FALSE}
qqnorm(Ref_DK$TrDur_us)
qqline(Ref_DK$TrDur_us)
ks.test(Ref_DK$TrDur_us, "pnorm", mean = mean(Ref_DK$TrDur_us), sd = sd(Ref_DK$TrDur_us))
#very small p-value (< 2.2e-16) suggests that the sample distribution significantly differs from the normal distribution
```

No noticeable normal distribution in Q-Q-Plot which is underlined by the result of the Kolmogorov-Smirnov-Test. <br>
The small p-value (<2.2e-16) suggest that the distribution significantly differs from normal distribution. <br>

Also checking for homogeneity of variances (Levenes Test):

```{r leveneTest 4, echo=FALSE}
leveneTest(TrDur_us ~ as.factor(c(rep(1, nrow(Ref_DE)), rep(2, nrow(Ref_DK)))), data = data.frame(TrDur_us = c(Ref_DE$TrDur_us, Ref_DK$TrDur_us)))
#p-value (<2.2e-16) < 0.05 --> statistically significant difference in variances between the groups
#null hypothesis of homogeneity of variances rejected
```

p-value (<2.2e-16) < 0.05 <br>
--> statistically significant difference in variances between the groups <br>
Meaning the null hypothesis of homogeneity of variances is rejected. 

## Statistical Test

```{r test 4, echo=FALSE}
#t.test(Ref_DE$TrDur_us, Ref_DK$TrDur_us, var.equal = FALSE)
wilcox.test(Ref_DE$TrDur_us, Ref_DK$TrDur_us)
#p (<2.2e-16) < 0.05 --> there is a significant difference between the datasets for the average Train Duration
```

p (<2.2e-16) < 0.05 <br>
--> there is a significant difference between the datasets for the average Train Duration <br>
Null Hypothesis rejected.

### Statistical Tests, pairwise (Mann-Whitney)

```{r wilcox DE, echo=FALSE}
wilcox.test(Ref_DE$MinICI_us, Ref_DE$medianKHz)
wilcox.test(Ref_DE$MinICI_us, Ref_DE$NofClstrs)
wilcox.test(Ref_DE$MinICI_us, Ref_DE$nActualClx)
wilcox.test(Ref_DE$MinICI_us, Ref_DE$AvPRF)
wilcox.test(Ref_DE$NofClstrs, Ref_DE$medianKHz)
wilcox.test(Ref_DE$nActualClx, Ref_DE$medianKHz)
wilcox.test(Ref_DE$AvPRF, Ref_DE$medianKHz)
wilcox.test(Ref_DE$NofClstrs, Ref_DE$nActualClx)
wilcox.test(Ref_DE$NofClstrs, Ref_DE$AvPRF)
wilcox.test(Ref_DE$nActualClx, Ref_DE$AvPRF)
```
```{r wilcox DK, echo=FALSE}
wilcox.test(Ref_DK$MinICI_us, Ref_DK$medianKHz)
wilcox.test(Ref_DK$MinICI_us, Ref_DK$NofClstrs)
wilcox.test(Ref_DK$MinICI_us, Ref_DK$nActualClx)
wilcox.test(Ref_DK$MinICI_us, Ref_DK$AvPRF)
wilcox.test(Ref_DK$NofClstrs, Ref_DK$medianKHz)
wilcox.test(Ref_DK$nActualClx, Ref_DK$medianKHz)
wilcox.test(Ref_DK$AvPRF, Ref_DK$medianKHz)
wilcox.test(Ref_DK$NofClstrs, Ref_DK$nActualClx)
wilcox.test(Ref_DK$NofClstrs, Ref_DK$AvPRF)
wilcox.test(Ref_DK$nActualClx, Ref_DK$AvPRF)
```

### Correlations

```{r cor, echo=FALSE}
pairs.panels(Ref_both[, c("MinICI_us", "medianKHz", "NofClstrs", "AvPRF", "nActualClx")], 
             method = "spearman", # Use Spearman correlation
             hist.col = "blue",   # Color of histograms
             density = TRUE,      # Add density plots
             ellipses = TRUE)     # Add correlation ellipses
```

```{r cor DE, echo=FALSE}
pairs.panels(Ref_DE[, c("MinICI_us", "medianKHz", "NofClstrs", "AvPRF", "nActualClx")], 
             method = "spearman", # Use Spearman correlation
             hist.col = "green",   # Color of histograms
             density = TRUE,      # Add density plots
             ellipses = TRUE)     # Add correlation ellipses
```

```{r cor DK, echo=FALSE}
pairs.panels(Ref_DK[, c("MinICI_us", "medianKHz", "NofClstrs", "AvPRF", "nActualClx")], 
             method = "spearman", # Use Spearman correlation
             hist.col = "orange",   # Color of histograms
             density = TRUE,      # Add density plots
             ellipses = TRUE)     # Add correlation ellipses
```